Impact of Roe v. Wade’s Overturn: Abortion Restrictions and Infant Mortality in the U.S. Using a Difference-in-Differences Analysis

### 2024-12-06 paper/ExaminingtheimpactOverturnofRoev.Wade:BanningofAbortiononInfantMortalityRatesintheUnitedStatesUsingaDifference-in-DifferencesApproach.qmd
fix gramma error

In this study, we utilized several R packages to streamline data
manipulation, modeling, and visualization. The `tidyverse` package
[@tidyverse] was central to our workflow, providing tools for data
wrangling and analysis, while `arrow` [@arrow] efficiently managed
parquet files for larger datasets. We used `modelsummary` package
[@modelsummary] to generate clean and interpretable summaries of our
regression models. We used `knitr` [@knitr] and `kableExtra`
[@kableExtra] to generate clean tables and reproducible reports, and
`usmap` for geographical visualizations. We use `dplyr` [@dplyr] to manipulate data. We use `testthat` [@testthat] to test the simulated and analysis dataset. We use `usmap` [@usmap] to plot and analyze U.S. state and county-level data on maps by providing preformatted map geometries and easy integration with data. We use `here` [@here] package simplifies the management of file paths, ensuring a reproducible
workflow by creating a consistent point of reference for file locations,
regardless of the operating system or working directory. Data cleaning
was facilitated by `janitor` [@janitor], while `lubridate` [@lubridate]
handled date-time variables. Visualization tasks were enhanced with
`ggplot2` [@ggplot2], and Bayesian regression modeling was conducted
using `rstanarm` [@rstanarm]. We use `fixest` [@fixest] to conduct the
event study. We use `caret` [@caret] to conduct the cross-validation of
the model. Together, these packages supported an efficient, reproducible
workflow for analyzing and presenting our findings.

@@

State (`state`) referes to the state of the mother’s legal residence at the time of birth. This
data is categorized by the mother’s state of residence and excludes territories.

@@

Define $y_i$ as as the infant death rate per 1,000 live births, and
$\gamma_j$ as the random state effect for state $j$. 
$$
\begin{aligned}
y_i|\mu_i,\sigma &\sim \mbox{Normal}(\mu_i, \sigma)\\
\text{First Model}: \mu_i &= \beta_0 + \beta_1\cdot\text{After Injunction} + \beta_2\cdot\text{Abortion Illegal}\\&+\beta_3\cdot\text{After Injunction}\cdot\text{Abortion Illegal}_i+\gamma_j\\
\text{Second Model}:\mu_i &= \beta_0 + \beta_1\cdot\text{After Injunction}_i + \beta_2\cdot\text{Abortion Illegal}_i\\&+ \beta_3\cdot\text{After Injunction}_i\cdot\text{Abortion Illegal}_i\\&+\beta_4\cdot\text{Year of Death}_i+\beta_5\cdot\text{Age of Mother}_i+\beta_6\cdot\text{Mother's Single Race}_i+\gamma_j\\
\beta_0 &\sim \mbox{Normal}(0, 2.5)\\
\beta_1 &\sim \mbox{Normal}(0, 2.5)\\
\beta_2 &\sim \mbox{Normal}(0, 2.5)\\
\beta_3 &\sim \mbox{Normal}(0, 2.5)\\
\beta_4 &\sim \mbox{Normal}(0, 2.5)\\
\beta_5 &\sim \mbox{Normal}(0, 2.5)\\
\beta_6 &\sim \mbox{Normal}(0, 2.5)\\
\gamma_j&\sim \text{Normal}(0, \sigma_j^2)\\
\sigma &\sim \mbox{Exponential}(1)
\end{aligned}
$$

>>>>

In this study, we utilized several R packages to streamline data
manipulation, modeling, and visualization. The `tidyverse` package
[@tidyverse] was central to our workflow, providing tools for data
wrangling and analysis, while `arrow` [@arrow] efficiently managed
parquet files for larger datasets. We used `modelsummary` package
[@modelsummary] to generate clean and interpretable summaries of our
regression models. We used `knitr` [@knitr] and `kableExtra`
[@kableExtra] to generate clean tables and reproducible reports, and
`usmap` for geographical visualizations. We use `dplyr` [@dplyr] to manipulate data. We use `testthat` [@testthat] to test the simulated and analysis dataset. We use `usmap` [@usmap] to plot and analyze U.S. state and county-level data on maps by providing preformatted map geometries and easy integration with data. We use `here` [@here] package to simplify the management of file paths, ensuring a reproducible
workflow by creating a consistent point of reference for file locations,
regardless of the operating system or working directory. Data cleaning
was facilitated by `janitor` [@janitor], while `lubridate` [@lubridate]
handled date-time variables. Visualization tasks were enhanced with
`ggplot2` [@ggplot2], and Bayesian regression modeling was conducted
using `rstanarm` [@rstanarm]. We use `fixest` [@fixest] to conduct the
event study. We use `caret` [@caret] to conduct the cross-validation of
the model. Together, these packages supported an efficient, reproducible
workflow for analyzing and presenting our findings.

@@

State (`state`) refers to the state of the mother’s legal residence at the time of birth. This
data is categorized by the mother’s state of residence and excludes territories.

@@

Define $y_i$ as the infant death rate per 1,000 live births, and
$\gamma_j$ as the random state effect for state $j$. 
$$
\begin{aligned}
y_i|\mu_i,\sigma &\sim \mbox{Normal}(\mu_i, \sigma)\\
\text{First Model}: \mu_i &= \beta_0 + \beta_1\cdot\text{After Injunction} + \beta_2\cdot\text{Abortion Illegal}\\&+\beta_3\cdot\text{After Injunction}\cdot\text{Abortion Illegal}_i+\gamma_j\\
\text{Second Model}:\mu_i &= \beta_0 + \beta_1\cdot\text{After Injunction}_i + \beta_2\cdot\text{Abortion Illegal}_i\\&+ \beta_3\cdot\text{After Injunction}_i\cdot\text{Abortion Illegal}_i\\&+\beta_4\cdot\text{Year of Death}_i+\beta_5\cdot\text{Age of Mother}_i+\beta_6\cdot\text{Mother's Single Race}_i+\gamma_j\\
\beta_0 &\sim \mbox{Normal}(0, 2.5)\\
\beta_1 &\sim \mbox{Normal}(0, 2.5)\\
\beta_2 &\sim \mbox{Normal}(0, 2.5)\\
\beta_3 &\sim \mbox{Normal}(0, 2.5)\\
\beta_4 &\sim \mbox{Normal}(0, 2.5)\\
\beta_5 &\sim \mbox{Normal}(0, 2.5)\\
\beta_6 &\sim \mbox{Normal}(0, 2.5)\\
\gamma_j&\sim \text{Normal}(0, \sigma_j^2)\\
\sigma &\sim \mbox{Exponential}(1)
\end{aligned}
$$
